%%
%% This is file `sample-acmsmall-conf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall-conf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall-conf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall,nonacm]{acmart}
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
%\pagestyle{plain}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
    
\definecolor{beaver}{RGB}{215, 63, 9}
\definecolor{luminance}{RGB}{255, 181, 0}


\usepackage{amsmath}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenc}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
%\setcopyright{none}
%\copyrightyear{}
%\acmYear{}
%\acmDOI{}

%% These commands are for a PROCEEDINGS abstract or paper.
%\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
%  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY}
%\acmPrice{15.00}
%\acmISBN{978-1-4503-XXXX-X/18/06}


%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Password Strength Meters}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{David Chen}
\email{chend2@oregonstate.edu}
\affiliation{%
  \institution{Oregon State University}
  \city{Corvallis}
  \state{Oregon}
  \country{USA}
  \postcode{97331}
}

\author{Abdullah Saydemir}
\email{saydemia@oregonstate.edu}
\affiliation{%
  \institution{Oregon State University}
  \city{Corvallis}
  \state{Oregon}
  \country{USA}
  \postcode{97331}
}

\author{Lindy Voss}
\email{vossli@oregonstate.edu}
\affiliation{%
  \institution{Oregon State University}
  \city{Corvallis}
  \state{Oregon}
  \country{USA}
  \postcode{97331}
}


%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Chen, Voss \& Saydemir}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    \textcolor{beaver}{[ Abstract will be written here when the text is completed ]}
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002978.10003029.10011703</concept_id>
       <concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10002991.10002992</concept_id>
       <concept_desc>Security and privacy~Authentication</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10003029</concept_id>
       <concept_desc>Security and privacy~Human and societal aspects of security and privacy</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10002979.10002983</concept_id>
       <concept_desc>Security and privacy~Cryptanalysis and other attacks</concept_desc>
       <concept_significance>100</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Usability in security and privacy}
\ccsdesc[300]{Security and privacy~Authentication}
\ccsdesc[500]{Security and privacy~Human and societal aspects of security and privacy}
\ccsdesc[100]{Security and privacy~Cryptanalysis and other attacks}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{security, password, memory}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
Passwords are used nowadays as the first line of defense to protect information along with other options \cite{ma_2010}. Though passwords are meant to be solid and secure, they have flaws that can be easily exploited. In 2019, nearly 1 terabyte of user account data including passwords previously collected from Yahoo and some other website breaches were for sale \cite{hunt_2019}. Hashes of mostly used passwords, including those from Yahoo!, get uploaded to haveibeenpwned.com for the public to check whether their passwords were breached. Among 700 million unique passwords, two of the most commonly used were “123456” (more than 24 million times) and “password” (more than 3 million times) which can be broken in virtually zero time.


To prevent these from happening, the websites benefit from password composition rules (e.g. password must contain both lowercase and uppercase letters). Studies show that enforcing these rules helps users to choose stronger passwords \cite{shay_2010}. However, this also increases user fatigue since the average number of website accounts a single user has is around 25 \cite{lee_2014,waugh_2012}  and continues to increase. This exhaustion leads to people creating one (or a few) strong passwords, often reusing those passwords across many websites \cite{das_2014} and preserving them for long periods \cite{bhagavatula_2020,florencio_2007}. Moreover, only one third of people change their password after a breach and most of them change their passwords to weaker or equally strong ones \cite{bhagavatula_2020}. This leads to large collections of passwords (that are gathered on the internet) being used in brute force attacks over and over again, often with some success due to all of the unchanged passwords. There are also highly specific versions of dictionaries used in these attacks that are categorized according to attack type, platform, website, language and even to geological location.


The problem with strong and secure passwords is nobody can remember them. Especially with the amount of websites and accounts people hold nowadays. People are told to create strong passwords and in order to do that they have to avoid short passwords, birthdays, names, places, dictionary words, and the list goes on \cite{lee_2014}. Basically you are not supposed to use anything remotely memorable. Say you come up with a strong password like Jf5ruEf392HQlx and somehow memorize it. Well you can only use that for one account and the rest all have to be different but just as strong \cite{rubenking_2021}. This begs the big question of how do you create a strong password (for each account) that is realistic to memorize for practical use? There are tools and different tricks that have been found to work to help people balance these two very important factors when deciding on a password so they can not only remember their passwords, but also keep their information safe.

In this article, we propose the following contributions:

\begin{itemize}
\item We will have a look at password entropy calculation to understand what dictates the strength of passwords.
\item We will introduce different password cracking techniques used to understand how passwords are being solved to better adapt new passwords to prevent cracking
\item We will introduce password creation techniques and guidelines, and discuss their strengths and weaknesses (or flaws) in relation to password strength meters.
\item We will introduce memorable passwords that are still strong to resist brute force / dictionary attacks.
\end{itemize}

In the next section, we will present an overview of concepts that are helpful to understand how a strong password is created.

\section{Concepts}

\subsection{Password Entropy / Strength}
Password entropy is a measurement how strong a password is. It is used to estimate effectiveness of a password against guessing or brute-force attacks. Higher entropy means stronger passwords whereas lower entropy means weaker. It depends on the complexity and length \cite{hu_2018}.\newline 

To put it in mathematical way, let's say \emph{P} is a password of length \emph{l} and consists of characters in alphabet $\Sigma$ size of which is equal to \emph{N}. That is, \emph{N} = |$\Sigma$| . Then, the entropy \emph{E} of this password \emph{P} is calculated with the following equation :
\begin{equation} \label{eq:1}
E = \log_2 N^l = l \cdot \log_2 N
\end{equation}

We can directly deduce that increasing either of \emph{L} or \emph{N} creates stronger passwords. Therefore, to create a stronger password user should either choose a larger set of characters or make the password longer. It is obvious that for some specific numbers, entropy of longer password from a smaller set can be similar to entropy of the shorter password from a larger set. For example, let's consider two passwords. $\emph{P}_1$ consists of digits only and is of length 11. $\emph{P}_2$ consists of lowercase Latin letters and is of length 8. If we calculate the entropy of both using set sizes given in \autoref{tab:sets} :
\begin{align*}\label{eq:p1p2calc}
    \emph{E}_1 = \log_2 10^{11} &= 11 \cdot \log_2 10 = 11 \times 3.32 &\emph{E}_1 \approx 36.5 \\
    \emph{E}_2 = \log_2 26^8  &= 8  \cdot \log_2 26   = 8 \times 4.70 &\emph{E}_2 \approx 37.6
\end{align*}

\noindent we can see that there is not much difference between the entropy of the two passwords. However, it is important to note that entropy itself is \textbf{NOT} all it matters. Consider the same passwords $\emph{P}_1$ and $\emph{P}_2$ but this time let $\emph{P}_1$ be the string "46583159027" and $\emph{P}_2$ be the string "password". Though they have similar entropy, $\emph{P}_2$ is extremely weak and it can be broken in virtually zero time since it appears in leaked password lists.
\begin{center}
\begin{table}[h!]

 \begin{tabular}[c]{| l  c  c |} 
\hline 
 Character Set & Elements & Set Size \\
 \hline\hline
 Digits & 0-9 & 10  \\ 
 \hline
 Latin letters (lowercase) & a-z & 26 \\
 \hline
 Latin letters (uppercase) & A-Z & 26  \\
 \hline
 Alphanumeric & a-z $\cup$  0-9 & 36  \\
 \hline
 Alphanumeric and Uppercase & a-z $\cup$ A-Z $\cup$ 0-9 & 62 \\  
 \hline
 Symbols (US keyboard) & $~!@\#\$\%\&*()-+=[]\{\}|;$ etc. & 32  \\
 \hline
 \end{tabular}
 \caption{Most Used Password Sets}
 \label{tab:sets}
\end{table}
\end{center}
Therefore, aside from the two parameters of \autoref{eq:1}, \emph{unpredictability}  is another factor that has an effect on password strength. We will talk about this concept in detail in \autoref{creation}.

\newpage
\subsection{Password Cracking-Techniques}

In order to know how to to best create a strong memorable password, it is important to know how many passwords are cracked. There are many different types of techniques that a hacker can use to figure out one's passwords. Each technique share the same end goal, but their execution differs.

\textcolor{beaver}{[ Generalized the methods into three categories: password-focused, human-focused, and computer-focused; this section will be focused on the password-focused, and human-focused ]}\newline


\textcolor{beaver}{[ Go into how each technique might affect a certain password; how effective some are towards certain types of passwords v.s. other types; Lead into the next few examples of other techniques besides brute forcing ]}

\subsubsection{Brute Force}

\textcolor{beaver}{[ What is brute forcing? How it is related to entropy? ]}

\subsubsection{Dictionary Attack/Mask Attack} \label{dictionary/mask}
\textcolor{beaver}{[ How this technique differs from brute forcing; What is the difference between Dictionary and Mask attacks? ]}


\section{Password Creation Techniques} \label{creation}

\textcolor{beaver}{[Quick maybe paragraph or 2 introduction to the section before going into the sub sections]}

\subsection{Strong Passwords}

\textcolor{beaver}{[ In this section we will analyze two password creation techniques ]}

\subsection{Memorable Passwords}

The easiest way to create a memorable password is to use something important to you or something that you like. Just using the names, dates, and other words is not secure though \cite{rubenking_2021}. The best password you can create is the one you cannot remember which is why this is such a dilemma \cite{lee_2014}. Human memory, as it turns out, is the biggest flaw of passwords \cite{yan_2000}. So we have to find a way to balance both of them in order to keep our information secure. There are multiple techniques and tools that can be used to help like using something important as we mentioned, or using outside management tools.

\textcolor{beaver}{[ This section will go on to describe some techniques to creating passwords would still be considered strong, but can be much more memorable as well as other tools that can help with managing strong passwords (1.5-2 pages) ]}


\section{Results \& Analysis}
\textcolor{beaver}{[ In this section we will briefly talk about the previous section and analyze the password creation technique we described, with examples ]} 

\section{Contributions}
\begin{itemize}
\item{\verb|Lindy Voss       |} : Introduction, Memorable Passwords, References
\item{\verb|Abdullah Saydemir|} : Introduction, Concepts, References
\item{\verb|David Chen       |} : Introduction, Concepts
\end{itemize}


\section{Conclusion}
\textcolor{beaver}{[ Conclusion to the report that we will probably do last so it sums everything up nicely and closes out the paper and everything was found/talked about ]}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{psm_bibliography}

%%
%% If your work has an appendix, this is the place to put it.
\appendix


\end{document}
\endinput
%%
%% End of file `password_strength.tex'.
